---
title: "EdgeR_all_files"
author: "Thanawit Ouioutai (n10739921)"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup library
```{r load_library}
library(tidyverse)
library(dplyr)
library(RColorBrewer)
library(DESeq2)
library(pheatmap)
library(DEGreport)
library(ggplot2)
library(ggrepel)
library(ggpubr)
library(edgeR)
```

## Reading in the data
```{r}
raw_data355 <- read_tsv("3_500_500.tsv", show_col_types = FALSE)
raw_data372 <- read_tsv("3_750_250.tsv", show_col_types = FALSE)
raw_data3100 <- read_tsv("3_1000_0.tsv", show_col_types = FALSE)

raw_data655 <- read_tsv("6_500_500.tsv", show_col_types = FALSE)
raw_data672 <- read_tsv("6_750_250.tsv", show_col_types = FALSE)
raw_data6100 <- read_tsv("6_1000_0.tsv", show_col_types = FALSE)

raw_data955 <- read_tsv("9_500_500.tsv", show_col_types = FALSE)
raw_data972 <- read_tsv("9_750_250.tsv", show_col_types = FALSE)
raw_data9100 <- read_tsv("9_1000_0.tsv", show_col_types = FALSE)
```

## Build a DGEList object
```{r}
# build a DGEList object from the rawdata count matrix
# "group" variable, factor giving the experimental condition for each sample.
group_condition_3 <- c("condition 1","condition 1","condition 1",
                       "condition 2","condition 2","condition 2")
group_condition_6 <- c("condition 1","condition 1","condition 1", 
                       "condition 1","condition 1","condition 1",
                       "condition 2","condition 2","condition 2",
                       "condition 2","condition 2","condition 2")  
group_condition_9 <- c("condition 1","condition 1","condition 1", 
                       "condition 1","condition 1","condition 1",
                       "condition 1","condition 1","condition 1",
                       "condition 2","condition 2","condition 2",
                       "condition 2","condition 2","condition 2",
                       "condition 2","condition 2","condition 2") 

gene_table_355 <- data.frame(gene = paste('g',rownames(raw_data355),sep=""),
                             stringsAsFactors = FALSE)
gene_table_372 <- data.frame(gene = paste('g',rownames(raw_data372),sep=""),
                             stringsAsFactors = FALSE)
gene_table_3100 <- data.frame(gene = paste('g',rownames(raw_data3100),sep=""),
                             stringsAsFactors = FALSE)

gene_table_655 <- data.frame(gene = paste('g',rownames(raw_data655),sep=""),
                             stringsAsFactors = FALSE)
gene_table_672 <- data.frame(gene = paste('g',rownames(raw_data672),sep=""),
                             stringsAsFactors = FALSE)
gene_table_6100 <- data.frame(gene = paste('g',rownames(raw_data6100),sep=""),
                             stringsAsFactors = FALSE)

gene_table_955 <- data.frame(gene = paste('g',rownames(raw_data955),sep=""),
                             stringsAsFactors = FALSE)
gene_table_972 <- data.frame(gene = paste('g',rownames(raw_data972),sep=""),
                             stringsAsFactors = FALSE)
gene_table_9100 <- data.frame(gene = paste('g',rownames(raw_data9100),sep=""),
                             stringsAsFactors = FALSE)


dge_355 <- DGEList(counts = raw_data355[, 2:7], group = group_condition_3, genes = gene_table_355)
dge_372 <- DGEList(counts = raw_data372[, 2:7], group = group_condition_3, genes = gene_table_372)
dge_3100 <- DGEList(counts = raw_data3100[, 2:7], group = group_condition_3, genes = gene_table_3100)

dge_655 <- DGEList(counts = raw_data655[, 2:13], group = group_condition_6, genes = gene_table_655)
dge_672 <- DGEList(counts = raw_data672[, 2:13], group = group_condition_6, genes = gene_table_672)
dge_6100 <- DGEList(counts = raw_data6100[, 2:13], group = group_condition_6, genes = gene_table_6100)

dge_955 <- DGEList(counts = raw_data955[, 2:19], group = group_condition_9, genes = gene_table_955)
dge_972 <- DGEList(counts = raw_data972[, 2:19], group = group_condition_9, genes = gene_table_972)
dge_9100 <- DGEList(counts = raw_data9100[, 2:19], group = group_condition_9, genes = gene_table_9100)
```

## Filtering
Genes with very low counts across all libraries provide little evidence for differential expression. These genes should be filtered out prior to further analysis.

```{r}
keep_355 <- filterByExpr(dge_355)
dge_355 <- dge_355[keep_355, , keep.lib.sizes=FALSE]
keep_372 <- filterByExpr(dge_372)
dge_372 <- dge_372[keep_372, , keep.lib.sizes=FALSE]
keep_3100 <- filterByExpr(dge_3100)
dge_3100 <- dge_3100[keep_3100, , keep.lib.sizes=FALSE]

keep_655 <- filterByExpr(dge_655)
dge_655 <- dge_655[keep_655, , keep.lib.sizes=FALSE]
keep_672 <- filterByExpr(dge_672)
dge_672 <- dge_672[keep_672, , keep.lib.sizes=FALSE]
keep_6100 <- filterByExpr(dge_6100)
dge_6100 <- dge_6100[keep_6100, , keep.lib.sizes=FALSE]

keep_955 <- filterByExpr(dge_955)
dge_955 <- dge_955[keep_955, , keep.lib.sizes=FALSE]
keep_972 <- filterByExpr(dge_972)
dge_972 <- dge_972[keep_972, , keep.lib.sizes=FALSE]
keep_9100 <- filterByExpr(dge_9100)
dge_9100 <- dge_9100[keep_9100, , keep.lib.sizes=FALSE]
```

## TMM normalization
Once a DGEList has been created, we calculate between-sample (TMM) normalization factors, using the calcNormFactors function in edgeR.

```{r}
# normalization
dge_355 <- calcNormFactors(dge_355)
dge_355$samples
dge_372 <- calcNormFactors(dge_372)
dge_372$samples
dge_3100 <- calcNormFactors(dge_3100)
dge_3100$samples

dge_655 <- calcNormFactors(dge_655)
dge_655$samples
dge_672 <- calcNormFactors(dge_672)
dge_672$samples
dge_6100 <- calcNormFactors(dge_6100)
dge_6100$samples

dge_955 <- calcNormFactors(dge_955)
dge_955$samples
dge_972 <- calcNormFactors(dge_972)
dge_972$samples
dge_9100 <- calcNormFactors(dge_9100)
dge_9100$samples
```

## Data exploration
An MDS plot shows the relative similarities of the samples.

edgeR contains a function plotMDS, which operates on a DGEList object and generates a two-dimensional MDS representation of the samples. The default distance between two samples can be interpreted as the “typical” log fold change between the two samples, for the genes that are most different between them

```{r}
# plotMDS
p355 <- plotMDS(dge_355, top = 1000, labels = NULL, 
                col = as.numeric(dge_355$samples$group), cex = 0.6)
#legend("bottomright", as.character(unique(dge$samples$group)), col = 1:2, pch = 19)
p372 <- plotMDS(dge_372, top = 1000, labels = NULL, 
                col = as.numeric(dge_372$samples$group), cex = 0.6)
p3100 <- plotMDS(dge_3100, top = 1000, labels = NULL, 
                col = as.numeric(dge_3100$samples$group), cex = 0.6)

p655 <- plotMDS(dge_655, top = 1000, labels = NULL, 
                col = as.numeric(dge_655$samples$group), cex = 0.6)
p672 <- plotMDS(dge_672, top = 1000, labels = NULL, 
                col = as.numeric(dge_672$samples$group), cex = 0.6)
p6100 <- plotMDS(dge_6100, top = 1000, labels = NULL, 
                col = as.numeric(dge_6100$samples$group), cex = 0.6)

p955 <- plotMDS(dge_955, top = 1000, labels = NULL, 
                col = as.numeric(dge_955$samples$group), cex = 0.6)
p972 <- plotMDS(dge_972, top = 1000, labels = NULL, 
                col = as.numeric(dge_972$samples$group), cex = 0.6)
p9100 <- plotMDS(dge_9100, top = 1000, labels = NULL, 
                col = as.numeric(dge_9100$samples$group), cex = 0.6)

figure <- ggarrange(p355, p372, p3100,
                    p655, p672, p6100,
                    p955, p972, p9100,
                    labels = c("3_500_500", "3_750_250", "3_1000_0", 
                               "6_500_500", "6_750_250", "6_1000_0", 
                               "9_500_500", "9_750_250", "9_1000_0"),
                    ncol = 3, nrow = 3)
annotate_figure(figure, top = text_grob("A two-dimensional MDS representation of the samples"))
annotate_figure
```

## Differential expression analysis
Next we will show how to perform differential expression analysis with edgeR. Recall that we have a DGEList (dge), containing all the necessary information:

### The classic edgeR pipeline: pairwise comparisons between two or more groups

- Estimating dispersions

- Testing for DE genes

For all the Next-Gen squencing data analyses we consider here, people are most interested in finding differentially expressed genes/tags between two (or more) groups. Once negative binomial models are fitted and dispersion estimates are obtained, we can proceed with testing procedures for determining differential expression using the exact test.

The exact test is based on the qCML methods. Knowing the conditional distribution for the sum of counts in a group, we can compute exact p-values by summing over all sums of counts that have a probability less than the probability under the null hypothesis of the observed sum of counts. The exact test for the negative binomial distribution has strong parallels with Fisher’s exact test.

As we dicussed in the previous section, the exact test is only applicable to experiments with a single factor. The testing can be done by using the function exactTest(), and the function allows both common dispersion and tagwise dispersion approaches.

```{r}
# Estimating dispersions and testing for DE genes

# _3
st_time_355 = Sys.time()
dge_355 <- estimateDisp(dge_355)
et_355 <- exactTest(dge_355)
end_time_355 = Sys.time()
edgeR_time_355 <- end_time_355 - st_time_355

st_time_372 = Sys.time()
dge_372 <- estimateDisp(dge_372)
et_372 <- exactTest(dge_372)
end_time_372 = Sys.time()
edgeR_time_372 <- end_time_372 - st_time_372

st_time_3100 = Sys.time()
dge_3100 <- estimateDisp(dge_3100)
et_3100 <- exactTest(dge_3100)
end_time_3100 = Sys.time()
edgeR_time_3100 <- end_time_3100 - st_time_3100

# _6
st_time_655 = Sys.time()
dge_655 <- estimateDisp(dge_655)
et_655 <- exactTest(dge_655)
end_time_655 = Sys.time()
edgeR_time_655 <- end_time_655 - st_time_655

st_time_672 = Sys.time()
dge_672 <- estimateDisp(dge_672)
et_672 <- exactTest(dge_672)
end_time_672 = Sys.time()
edgeR_time_672 <- end_time_672 - st_time_672

st_time_6100 = Sys.time()
dge_6100 <- estimateDisp(dge_6100)
et_6100 <- exactTest(dge_6100)
end_time_6100 = Sys.time()
edgeR_time_6100 <- end_time_6100 - st_time_6100

# _9
st_time_955 = Sys.time()
dge_955 <- estimateDisp(dge_955)
et_955 <- exactTest(dge_955)
end_time_955 = Sys.time()
edgeR_time_955 <- end_time_955 - st_time_955

st_time_972 = Sys.time()
dge_972 <- estimateDisp(dge_972)
et_972 <- exactTest(dge_972)
end_time_972 = Sys.time()
edgeR_time_972 <- end_time_972 - st_time_972

st_time_9100 = Sys.time()
dge_9100 <- estimateDisp(dge_9100)
et_9100 <- exactTest(dge_9100)
end_time_9100 = Sys.time()
edgeR_time_9100 <- end_time_9100 - st_time_9100

# plot BCV
plotBCV(dge_355)
plotBCV(dge_372)
plotBCV(dge_3100)

plotBCV(dge_655)
plotBCV(dge_672)
plotBCV(dge_6100)

plotBCV(dge_955)
plotBCV(dge_972)
plotBCV(dge_9100)
```

```{r}
# topTags()
df_export_355 <- topTags(et_355, n = 1000, p.value = 0.05)
df_export_372 <- topTags(et_372, n = 1000, p.value = 0.05)
df_export_3100 <- topTags(et_3100, n = 1000, p.value = 0.05)

df_export_655 <- topTags(et_655, n = 1000, p.value = 0.05)
df_export_672 <- topTags(et_672, n = 1000, p.value = 0.05)
df_export_6100 <- topTags(et_6100, n = 1000, p.value = 0.05)

df_export_955 <- topTags(et_955, n = 1000, p.value = 0.05)
df_export_972 <- topTags(et_972, n = 1000, p.value = 0.05)
df_export_9100 <- topTags(et_9100, n = 1000, p.value = 0.05)

# export .csv files
write.csv(df_export_355,"V2_EdgeR_3_500_500.csv", row.names = FALSE)
write.csv(df_export_372,"V2_EdgeR_3_750_250.csv", row.names = FALSE)
write.csv(df_export_3100,"V2_EdgeR_3_1000_0.csv", row.names = FALSE)

write.csv(df_export_655,"V2_EdgeR_6_500_500.csv", row.names = FALSE)
write.csv(df_export_672,"V2_EdgeR_6_750_250.csv", row.names = FALSE)
write.csv(df_export_6100,"V2_EdgeR_6_1000_0.csv", row.names = FALSE)

write.csv(df_export_955,"V2_EdgeR_9_500_500.csv", row.names = FALSE)
write.csv(df_export_972,"V2_EdgeR_9_750_250.csv", row.names = FALSE)
write.csv(df_export_9100,"V2_EdgeR_9_1000_0.csv", row.names = FALSE)

# combind table
et_355$table <- cbind(et_355$genes,et_355$table)
et_372$table <- cbind(et_372$genes,et_372$table)
et_3100$table <- cbind(et_3100$genes,et_3100$table)

et_655$table <- cbind(et_655$genes,et_655$table)
et_672$table <- cbind(et_672$genes,et_672$table)
et_6100$table <- cbind(et_6100$genes,et_6100$table)

et_955$table <- cbind(et_955$genes,et_955$table)
et_972$table <- cbind(et_972$genes,et_972$table)
et_9100$table <- cbind(et_9100$genes,et_9100$table)
```

The total number of differentially expressed genes at 5% FDR is given by:
```{r}
# default: decideTests(object, adjust.method="BH", p.value = 0.05, lfc = 0, ...)
result_decideTest_355 <- decideTests(et_355, lfc = 0.58)
result_decideTest_372 <- decideTests(et_372, lfc = 0.58)
result_decideTest_3100 <- decideTests(et_3100, lfc = 0.58)

result_decideTest_655 <- decideTests(et_655, lfc = 0.58)
result_decideTest_672 <- decideTests(et_672, lfc = 0.58)
result_decideTest_6100 <- decideTests(et_6100, lfc = 0.58)

result_decideTest_955 <- decideTests(et_955, lfc = 0.58)
result_decideTest_972 <- decideTests(et_972, lfc = 0.58)
result_decideTest_9100 <- decideTests(et_9100, lfc = 0.58)

#summary(result_decideTest_655)
```

```{r}
# add the new column, which is the output of decideTest, into genes dataframe
et_355$table <- cbind(et_355$table,data.frame(result_decideTest_355@.Data))
# add the diffexpressed collumn
et_355$table$diffexpressed <- "NotSig"
# if condition.2.condition.1 == -1 set as "Down"
et_355$table$diffexpressed[et_355$table$condition.2.condition.1 == -1] <- "Down"
# if condition.2.condition.1 == 1 set as "Up" 
et_355$table$diffexpressed[et_355$table$condition.2.condition.1 == 1] <- "Up"
# if condition.2.condition.1 == 0 set as "NotSig" 
et_355$table$diffexpressed[et_355$table$condition.2.condition.1 == 0] <- "NotSig"
#head(et$table)

et_372$table <- cbind(et_372$table,data.frame(result_decideTest_372@.Data))
et_372$table$diffexpressed <- "NotSig"
et_372$table$diffexpressed[et_372$table$condition.2.condition.1 == -1] <- "Down"
et_372$table$diffexpressed[et_372$table$condition.2.condition.1 == 1] <- "Up"
et_372$table$diffexpressed[et_372$table$condition.2.condition.1 == 0] <- "NotSig"

et_3100$table <- cbind(et_3100$table,data.frame(result_decideTest_3100@.Data))
et_3100$table$diffexpressed <- "NotSig"
et_3100$table$diffexpressed[et_3100$table$condition.2.condition.1 == -1] <- "Down"
et_3100$table$diffexpressed[et_3100$table$condition.2.condition.1 == 1] <- "Up"
et_3100$table$diffexpressed[et_3100$table$condition.2.condition.1 == 0] <- "NotSig"

# _6
et_655$table <- cbind(et_655$table,data.frame(result_decideTest_655@.Data))
et_655$table$diffexpressed <- "NotSig"
et_655$table$diffexpressed[et_655$table$condition.2.condition.1 == -1] <- "Down"
et_655$table$diffexpressed[et_655$table$condition.2.condition.1 == 1] <- "Up"
et_655$table$diffexpressed[et_655$table$condition.2.condition.1 == 0] <- "NotSig"

et_672$table <- cbind(et_672$table,data.frame(result_decideTest_672@.Data))
et_672$table$diffexpressed <- "NotSig"
et_672$table$diffexpressed[et_672$table$condition.2.condition.1 == -1] <- "Down"
et_672$table$diffexpressed[et_672$table$condition.2.condition.1 == 1] <- "Up"
et_672$table$diffexpressed[et_672$table$condition.2.condition.1 == 0] <- "NotSig"

et_6100$table <- cbind(et_6100$table,data.frame(result_decideTest_6100@.Data))
et_6100$table$diffexpressed <- "NotSig"
et_6100$table$diffexpressed[et_6100$table$condition.2.condition.1 == -1] <- "Down"
et_6100$table$diffexpressed[et_6100$table$condition.2.condition.1 == 1] <- "Up"
et_6100$table$diffexpressed[et_6100$table$condition.2.condition.1 == 0] <- "NotSig"

# _9
et_955$table <- cbind(et_955$table,data.frame(result_decideTest_955@.Data))
et_955$table$diffexpressed <- "NotSig"
et_955$table$diffexpressed[et_955$table$condition.2.condition.1 == -1] <- "Down"
et_955$table$diffexpressed[et_955$table$condition.2.condition.1 == 1] <- "Up"
et_955$table$diffexpressed[et_955$table$condition.2.condition.1 == 0] <- "NotSig"

et_972$table <- cbind(et_972$table,data.frame(result_decideTest_972@.Data))
et_972$table$diffexpressed <- "NotSig"
et_972$table$diffexpressed[et_972$table$condition.2.condition.1 == -1] <- "Down"
et_972$table$diffexpressed[et_972$table$condition.2.condition.1 == 1] <- "Up"
et_972$table$diffexpressed[et_972$table$condition.2.condition.1 == 0] <- "NotSig"

et_9100$table <- cbind(et_9100$table,data.frame(result_decideTest_9100@.Data))
et_9100$table$diffexpressed <- "NotSig"
et_9100$table$diffexpressed[et_9100$table$condition.2.condition.1 == -1] <- "Down"
et_9100$table$diffexpressed[et_9100$table$condition.2.condition.1 == 1] <- "Up"
et_9100$table$diffexpressed[et_9100$table$condition.2.condition.1 == 0] <- "NotSig"
```

Plot log-fold change against log-counts per million, with DE genes highlighted:
```{r}
#plotMD(et)
#abline(h=c(-1, 1), col="blue")
```

The blue lines indicate 2-fold changes.

### Volcano plot
```{r}
arrange_et_355 <- et_355
arrange_et_372 <- et_372
arrange_et_3100 <- et_3100

arrange_et_655 <- et_655
arrange_et_672 <- et_672
arrange_et_6100 <- et_6100

arrange_et_955 <- et_955
arrange_et_972 <- et_972
arrange_et_9100 <- et_9100

# arrange the et by using p-value (ascending order). Actually, we identify the differential expression genes by using FDR or adjusted p-value.
arrange_et_355$table <- arrange_et_355$table %>% arrange(PValue) %>% mutate(genelabels = "")
arrange_et_355$table$genelabels[1:10] <- arrange_et_355$table$gene[1:10]
arrange_et_372$table <- arrange_et_372$table %>% arrange(PValue) %>% mutate(genelabels = "")
arrange_et_372$table$genelabels[1:10] <- arrange_et_372$table$gene[1:10]
arrange_et_3100$table <- arrange_et_3100$table %>% arrange(PValue) %>% mutate(genelabels = "")
arrange_et_3100$table$genelabels[1:10] <- arrange_et_3100$table$gene[1:10]

arrange_et_655$table <- arrange_et_655$table %>% arrange(PValue) %>% mutate(genelabels = "")
arrange_et_655$table$genelabels[1:10] <- arrange_et_655$table$gene[1:10]
arrange_et_672$table <- arrange_et_672$table %>% arrange(PValue) %>% mutate(genelabels = "")
arrange_et_672$table$genelabels[1:10] <- arrange_et_672$table$gene[1:10]
arrange_et_6100$table <- arrange_et_6100$table %>% arrange(PValue) %>% mutate(genelabels = "")
arrange_et_6100$table$genelabels[1:10] <- arrange_et_6100$table$gene[1:10]

arrange_et_955$table <- arrange_et_955$table %>% arrange(PValue) %>% mutate(genelabels = "")
arrange_et_955$table$genelabels[1:10] <- arrange_et_955$table$gene[1:10]
arrange_et_972$table <- arrange_et_972$table %>% arrange(PValue) %>% mutate(genelabels = "")
arrange_et_972$table$genelabels[1:10] <- arrange_et_972$table$gene[1:10]
arrange_et_9100$table <- arrange_et_9100$table %>% arrange(PValue) %>% mutate(genelabels = "")
arrange_et_9100$table$genelabels[1:10] <- arrange_et_9100$table$gene[1:10]
```

```{r}
# _3
volca_355 <- ggplot(data=arrange_et_355$table, aes(x=arrange_et_355$table$logFC, 
                          y=-log10(arrange_et_355$table$PValue))) + 
  geom_point(aes(col=diffexpressed)) + 
  theme_minimal() + 
  geom_text_repel(aes(label=genelabels), max.overlaps = getOption("ggrepel.max.overlaps", default = 10)) + 
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_vline(xintercept=c(0, 0), col="red") +
  geom_hline(yintercept=-log10(1.377256*10^-3), col="red") +
  ggtitle("3_500_500 differential expression") +
  xlab("log2 fold change") +
  ylab("-log10 p-value")

volca_372 <- ggplot(data=arrange_et_372$table, aes(x=arrange_et_372$table$logFC, 
                          y=-log10(arrange_et_372$table$PValue))) + 
  geom_point(aes(col=diffexpressed)) + 
  theme_minimal() + 
  geom_text_repel(aes(label=genelabels), max.overlaps = getOption("ggrepel.max.overlaps", default = 10)) + 
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_vline(xintercept=c(0, 0), col="red") +
  geom_hline(yintercept=-log10(1.498174*10^-3), col="red") +
  ggtitle("3_750_250 differential expression") +
  xlab("log2 fold change") +
  ylab("-log10 p-value")

volca_3100 <- ggplot(data=arrange_et_3100$table, aes(x=arrange_et_3100$table$logFC, 
                          y=-log10(arrange_et_3100$table$PValue))) + 
  geom_point(aes(col=diffexpressed)) + 
  theme_minimal() + 
  geom_text_repel(aes(label=genelabels), max.overlaps = getOption("ggrepel.max.overlaps", default = 10)) + 
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_vline(xintercept=c(0, 0), col="red") +
  geom_hline(yintercept=-log10(1.590792*10^-3), col="red") +
  ggtitle("3_1000_0 differential expression") +
  xlab("log2 fold change") +
  ylab("-log10 p-value")

# _6
volca_655 <- ggplot(data=arrange_et_655$table, aes(x=arrange_et_655$table$logFC, 
                          y=-log10(arrange_et_655$table$PValue))) + 
  geom_point(aes(col=diffexpressed)) + 
  theme_minimal() + 
  geom_text_repel(aes(label=genelabels), max.overlaps = getOption("ggrepel.max.overlaps", default = 10)) + 
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_vline(xintercept=c(0, 0), col="red") +
  geom_hline(yintercept=-log10(2.777723*10^-3), col="red") +
  ggtitle("6_500_500 differential expression") +
  xlab("log2 fold change") +
  ylab("-log10 p-value")

volca_672 <- ggplot(data=arrange_et_672$table, aes(x=arrange_et_672$table$logFC, 
                          y=-log10(arrange_et_672$table$PValue))) + 
  geom_point(aes(col=diffexpressed)) + 
  theme_minimal() + 
  geom_text_repel(aes(label=genelabels), max.overlaps = getOption("ggrepel.max.overlaps", default = 10)) + 
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_vline(xintercept=c(0, 0), col="red") +
  geom_hline(yintercept=-log10(2.953149*10^-3), col="red") +
  ggtitle("6_750_250 differential expression") +
  xlab("log2 fold change") +
  ylab("-log10 p-value")

volca_6100 <- ggplot(data=arrange_et_6100$table, aes(x=arrange_et_6100$table$logFC, 
                          y=-log10(arrange_et_6100$table$PValue))) + 
  geom_point(aes(col=diffexpressed)) + 
  theme_minimal() + 
  geom_text_repel(aes(label=genelabels), max.overlaps = getOption("ggrepel.max.overlaps", default = 10)) + 
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_vline(xintercept=c(0, 0), col="red") +
  geom_hline(yintercept=-log10(2.922512*10^-3), col="red") +
  ggtitle("6_1000_0 differential expression") +
  xlab("log2 fold change") +
  ylab("-log10 p-value")

#_9
volca_955 <- ggplot(data=arrange_et_955$table, aes(x=arrange_et_955$table$logFC, 
                          y=-log10(arrange_et_955$table$PValue))) + 
  geom_point(aes(col=diffexpressed)) + 
  theme_minimal() + 
  geom_text_repel(aes(label=genelabels), max.overlaps = getOption("ggrepel.max.overlaps", default = 10)) + 
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_vline(xintercept=c(0, 0), col="red") +
  geom_hline(yintercept=-log10(3.557585*10^-3), col="red") +
  ggtitle("9_500_500 differential expression") +
  xlab("log2 fold change") +
  ylab("-log10 p-value")

volca_972 <- ggplot(data=arrange_et_972$table, aes(x=arrange_et_972$table$logFC, 
                          y=-log10(arrange_et_972$table$PValue))) + 
  geom_point(aes(col=diffexpressed)) + 
  theme_minimal() + 
  geom_text_repel(aes(label=genelabels), max.overlaps = getOption("ggrepel.max.overlaps", default = 10)) + 
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_vline(xintercept=c(0, 0), col="red") +
  geom_hline(yintercept=-log10(3.600677*10^-3), col="red") +
  ggtitle("9_750_250 differential expression") +
  xlab("log2 fold change") +
  ylab("-log10 p-value")

volca_9100 <- ggplot(data=arrange_et_9100$table, aes(x=arrange_et_9100$table$logFC, 
                          y=-log10(arrange_et_9100$table$PValue))) + 
  geom_point(aes(col=diffexpressed)) + 
  theme_minimal() + 
  geom_text_repel(aes(label=genelabels), max.overlaps = getOption("ggrepel.max.overlaps", default = 10)) + 
  scale_color_manual(values=c("blue", "black", "red")) +
  geom_vline(xintercept=c(0, 0), col="red") +
  geom_hline(yintercept=-log10(3.322350*10^-3), col="red") +
  ggtitle("9_1000_0 differential expression") +
  xlab("log2 fold change") +
  ylab("-log10 p-value")

figure2 <- ggarrange(volca_355, volca_372, volca_3100,
                    volca_655, volca_672, volca_6100,
                    volca_955, volca_972, volca_9100,
                    labels = c("A", "B", "C", 
                               "D", "E", "F",
                               "G", "H", "I"),
                    ncol = 3, nrow = 3)
#figure2
```

This volcano plot represents differential expression of genes. On the x-axis we typically find the fold change and on the y-axis the p-value.

## Computational time EdgeR
```{r}
# present computational time
print(paste0("3_500_500 computational time: ", edgeR_time_355))
print(paste0("3_750_250 computational time: ", edgeR_time_372))
print(paste0("3_1000_0 computational time: ", edgeR_time_3100))

print(paste0("6_500_500 computational time: ", edgeR_time_655))
print(paste0("6_750_250 computational time: ", edgeR_time_672))
print(paste0("6_1000_0 computational time: ", edgeR_time_6100))

print(paste0("9_500_500 computational time: ", edgeR_time_955))
print(paste0("9_750_250 computational time: ", edgeR_time_972))
print(paste0("9_1000_0 computational time: ", edgeR_time_9100))
```

## Comparing the results with meta files

```{r}
meta_data355 <- read_tsv("3_500_500_meta.tsv", show_col_types = FALSE)
meta_data372 <- read_tsv("3_750_250_meta.tsv", show_col_types = FALSE)
meta_data3100 <- read_tsv("3_1000_0_meta.tsv", show_col_types = FALSE)

meta_data655 <- read_tsv("6_500_500_meta.tsv", show_col_types = FALSE)
meta_data672 <- read_tsv("6_750_250_meta.tsv", show_col_types = FALSE)
meta_data6100 <- read_tsv("6_1000_0_meta.tsv", show_col_types = FALSE)

meta_data955 <- read_tsv("9_500_500_meta.tsv", show_col_types = FALSE)
meta_data972 <- read_tsv("9_750_250_meta.tsv", show_col_types = FALSE)
meta_data9100 <- read_tsv("9_1000_0_meta.tsv", show_col_types = FALSE)
```

```{r}
# upregulated
# _3
meta355_up <- as.data.frame(meta_data355[meta_data355$upregulation == 1, 1])
edgeR_355_up <- as.data.frame(et_355$table[et_355$table$condition.2.condition.1 == 1, 1])
meta372_up <- as.data.frame(meta_data372[meta_data372$upregulation == 1, 1])
edgeR_372_up <- as.data.frame(et_372$table[et_372$table$condition.2.condition.1 == 1, 1])
meta3100_up <- as.data.frame(meta_data3100[meta_data3100$upregulation == 1, 1])
edgeR_3100_up <- as.data.frame(et_3100$table[et_3100$table$condition.2.condition.1 == 1, 1])

# _6
meta655_up <- as.data.frame(meta_data655[meta_data655$upregulation == 1, 1])
edgeR_655_up <- as.data.frame(et_655$table[et_655$table$condition.2.condition.1 == 1, 1])
meta672_up <- as.data.frame(meta_data672[meta_data672$upregulation == 1, 1])
edgeR_672_up <- as.data.frame(et_672$table[et_672$table$condition.2.condition.1 == 1, 1])
meta6100_up <- as.data.frame(meta_data6100[meta_data6100$upregulation == 1, 1])
edgeR_6100_up <- as.data.frame(et_6100$table[et_6100$table$condition.2.condition.1 == 1, 1])

# _9
meta955_up <- as.data.frame(meta_data955[meta_data955$upregulation == 1, 1])
edgeR_955_up <- as.data.frame(et_955$table[et_955$table$condition.2.condition.1 == 1, 1])
meta972_up <- as.data.frame(meta_data972[meta_data972$upregulation == 1, 1])
edgeR_972_up <- as.data.frame(et_972$table[et_972$table$condition.2.condition.1 == 1, 1])
meta9100_up <- as.data.frame(meta_data9100[meta_data9100$upregulation == 1, 1])
edgeR_9100_up <- as.data.frame(et_9100$table[et_9100$table$condition.2.condition.1 == 1, 1])

# change column name
# _3
colnames(meta355_up)[1] <- "genes"
colnames(edgeR_355_up)[1] <- "genes"
colnames(meta372_up)[1] <- "genes"
colnames(edgeR_372_up)[1] <- "genes"
colnames(meta3100_up)[1] <- "genes"
colnames(edgeR_3100_up)[1] <- "genes"

# _6
colnames(meta655_up)[1] <- "genes"
colnames(edgeR_655_up)[1] <- "genes"
colnames(meta672_up)[1] <- "genes"
colnames(edgeR_672_up)[1] <- "genes"
colnames(meta6100_up)[1] <- "genes"
colnames(edgeR_6100_up)[1] <- "genes"

# _9
colnames(meta955_up)[1] <- "genes"
colnames(edgeR_955_up)[1] <- "genes"
colnames(meta972_up)[1] <- "genes"
colnames(edgeR_972_up)[1] <- "genes"
colnames(meta9100_up)[1] <- "genes"
colnames(edgeR_9100_up)[1] <- "genes"

# find intersection genes
# _3
intersect_355_up <- intersect(meta355_up$genes, edgeR_355_up$genes)
intersect_372_up <- intersect(meta372_up$genes, edgeR_372_up$genes)
intersect_3100_up <- intersect(meta3100_up$genes, edgeR_3100_up$genes)

# _6
intersect_655_up <- intersect(meta655_up$genes, edgeR_655_up$genes)
intersect_672_up <- intersect(meta672_up$genes, edgeR_672_up$genes)
intersect_6100_up <- intersect(meta6100_up$genes, edgeR_6100_up$genes)

# _9
intersect_955_up <- intersect(meta955_up$genes, edgeR_955_up$genes)
intersect_972_up <- intersect(meta972_up$genes, edgeR_972_up$genes)
intersect_9100_up <- intersect(meta9100_up$genes, edgeR_9100_up$genes)
```

```{r}
# down-regulated
# _3
meta355_down <- as.data.frame(meta_data355[meta_data355$downregulation == 1, 1])
edgeR_355_down <- as.data.frame(et_355$table[et_355$table$condition.2.condition.1 == -1, 1])
meta372_down <- as.data.frame(meta_data372[meta_data372$downregulation == 1, 1])
edgeR_372_down <- as.data.frame(et_372$table[et_372$table$condition.2.condition.1 == -1, 1])
meta3100_down <- as.data.frame(meta_data3100[meta_data3100$downregulation == 1, 1])
edgeR_3100_down <- as.data.frame(et_3100$table[et_3100$table$condition.2.condition.1 == -1, 1])

# _6
meta655_down <- as.data.frame(meta_data655[meta_data655$downregulation == 1, 1])
edgeR_655_down <- as.data.frame(et_655$table[et_655$table$condition.2.condition.1 == -1, 1])
meta672_down <- as.data.frame(meta_data672[meta_data672$downregulation == 1, 1])
edgeR_672_down <- as.data.frame(et_672$table[et_672$table$condition.2.condition.1 == -1, 1])
meta6100_down <- as.data.frame(meta_data6100[meta_data6100$downregulation == 1, 1])
edgeR_6100_down <- as.data.frame(et_6100$table[et_6100$table$condition.2.condition.1 == -1, 1])

# _9
meta955_down <- as.data.frame(meta_data955[meta_data955$downregulation == 1, 1])
edgeR_955_down <- as.data.frame(et_955$table[et_955$table$condition.2.condition.1 == -1, 1])
meta972_down <- as.data.frame(meta_data972[meta_data972$downregulation == 1, 1])
edgeR_972_down <- as.data.frame(et_972$table[et_972$table$condition.2.condition.1 == -1, 1])
meta9100_down <- as.data.frame(meta_data9100[meta_data9100$downregulation == 1, 1])
edgeR_9100_down <- as.data.frame(et_9100$table[et_9100$table$condition.2.condition.1 == -1, 1])

# change column name
# _3
colnames(meta355_down)[1] <- "genes"
colnames(edgeR_355_down)[1] <- "genes"
colnames(meta372_down)[1] <- "genes"
colnames(edgeR_372_down)[1] <- "genes"
colnames(meta3100_down)[1] <- "genes"
colnames(edgeR_3100_down)[1] <- "genes"

# _6
colnames(meta655_down)[1] <- "genes"
colnames(edgeR_655_down)[1] <- "genes"
colnames(meta672_down)[1] <- "genes"
colnames(edgeR_672_down)[1] <- "genes"
colnames(meta6100_down)[1] <- "genes"
colnames(edgeR_6100_down)[1] <- "genes"

# _9
colnames(meta955_down)[1] <- "genes"
colnames(edgeR_955_down)[1] <- "genes"
colnames(meta972_down)[1] <- "genes"
colnames(edgeR_972_down)[1] <- "genes"
colnames(meta9100_down)[1] <- "genes"
colnames(edgeR_9100_down)[1] <- "genes"

# find intersection genes
# _3
intersect_355_down <- intersect(meta355_down$genes, edgeR_355_down$genes)
intersect_372_down <- intersect(meta372_down$genes, edgeR_372_down$genes)
intersect_3100_down <- intersect(meta3100_down$genes, edgeR_3100_down$genes)

# _6
intersect_655_down <- intersect(meta655_down$genes, edgeR_655_down$genes)
intersect_672_down <- intersect(meta672_down$genes, edgeR_672_down$genes)
intersect_6100_down <- intersect(meta6100_down$genes, edgeR_6100_down$genes)

# _9
intersect_955_down <- intersect(meta955_down$genes, edgeR_955_down$genes)
intersect_972_down <- intersect(meta972_down$genes, edgeR_972_down$genes)
intersect_9100_down <- intersect(meta9100_down$genes, edgeR_9100_down$genes)
```

```{r}
# DE
# _3
meta355_DE <- as.data.frame(meta_data355[meta_data355$differential.expression == 1, 1])
edgeR_355_DE <- as.data.frame(et_355$table[et_355$table$condition.2.condition.1 != 0, 1])
meta372_DE <- as.data.frame(meta_data372[meta_data372$differential.expression == 1, 1])
edgeR_372_DE <- as.data.frame(et_372$table[et_372$table$condition.2.condition.1 != 0, 1])
meta3100_DE <- as.data.frame(meta_data3100[meta_data3100$differential.expression == 1, 1])
edgeR_3100_DE <- as.data.frame(et_3100$table[et_3100$table$condition.2.condition.1 != 0, 1])

# _6
meta655_DE <- as.data.frame(meta_data655[meta_data655$differential.expression == 1, 1])
edgeR_655_DE <- as.data.frame(et_655$table[et_655$table$condition.2.condition.1 != 0, 1])
meta672_DE <- as.data.frame(meta_data672[meta_data672$differential.expression == 1, 1])
edgeR_672_DE <- as.data.frame(et_672$table[et_672$table$condition.2.condition.1 != 0, 1])
meta6100_DE <- as.data.frame(meta_data6100[meta_data6100$differential.expression == 1, 1])
edgeR_6100_DE <- as.data.frame(et_6100$table[et_6100$table$condition.2.condition.1 != 0, 1])

# _9
meta955_DE <- as.data.frame(meta_data955[meta_data955$differential.expression == 1, 1])
edgeR_955_DE <- as.data.frame(et_955$table[et_955$table$condition.2.condition.1 != 0, 1])
meta972_DE <- as.data.frame(meta_data972[meta_data972$differential.expression == 1, 1])
edgeR_972_DE <- as.data.frame(et_972$table[et_972$table$condition.2.condition.1 != 0, 1])
meta9100_DE <- as.data.frame(meta_data9100[meta_data9100$differential.expression == 1, 1])
edgeR_9100_DE <- as.data.frame(et_9100$table[et_9100$table$condition.2.condition.1 != 0, 1])

# change column name
# _3
colnames(meta355_DE)[1] <- "genes"
colnames(edgeR_355_DE)[1] <- "genes"
colnames(meta372_DE)[1] <- "genes"
colnames(edgeR_372_DE)[1] <- "genes"
colnames(meta3100_DE)[1] <- "genes"
colnames(edgeR_3100_DE)[1] <- "genes"

# _6
colnames(meta655_DE)[1] <- "genes"
colnames(edgeR_655_DE)[1] <- "genes"
colnames(meta672_DE)[1] <- "genes"
colnames(edgeR_672_DE)[1] <- "genes"
colnames(meta6100_DE)[1] <- "genes"
colnames(edgeR_6100_DE)[1] <- "genes"

# _9
colnames(meta955_DE)[1] <- "genes"
colnames(edgeR_955_DE)[1] <- "genes"
colnames(meta972_DE)[1] <- "genes"
colnames(edgeR_972_DE)[1] <- "genes"
colnames(meta9100_DE)[1] <- "genes"
colnames(edgeR_9100_DE)[1] <- "genes"

# find intersection genes
# _3
intersect_355_DE <- intersect(meta355_DE$genes, edgeR_355_DE$genes)
intersect_372_DE <- intersect(meta372_DE$genes, edgeR_372_DE$genes)
intersect_3100_DE <- intersect(meta3100_DE$genes, edgeR_3100_DE$genes)

# _6
intersect_655_DE <- intersect(meta655_DE$genes, edgeR_655_DE$genes)
intersect_672_DE <- intersect(meta672_DE$genes, edgeR_672_DE$genes)
intersect_6100_DE <- intersect(meta6100_DE$genes, edgeR_6100_DE$genes)

# _9
intersect_955_DE <- intersect(meta955_DE$genes, edgeR_955_DE$genes)
intersect_972_DE <- intersect(meta972_DE$genes, edgeR_972_DE$genes)
intersect_9100_DE <- intersect(meta9100_DE$genes, edgeR_9100_DE$genes)
```

```{r}
# acc_up
acc_355_up <- length(intersect_355_up)/nrow(meta355_up)
acc_372_up <- length(intersect_372_up)/nrow(meta372_up)
acc_3100_up <- length(intersect_3100_up)/nrow(meta3100_up)

acc_655_up <- length(intersect_655_up)/nrow(meta655_up)
acc_672_up <- length(intersect_672_up)/nrow(meta672_up)
acc_6100_up <- length(intersect_6100_up)/nrow(meta6100_up)

acc_955_up <- length(intersect_955_up)/nrow(meta955_up)
acc_972_up <- length(intersect_972_up)/nrow(meta972_up)
acc_9100_up <- length(intersect_9100_up)/nrow(meta9100_up)

# acc_down
acc_355_down <- length(intersect_355_down)/nrow(meta355_down)
acc_372_down <- length(intersect_372_down)/nrow(meta372_down)
acc_3100_down <- length(intersect_3100_down)/nrow(meta3100_down)

acc_655_down <- length(intersect_655_down)/nrow(meta655_down)
acc_672_down <- length(intersect_672_down)/nrow(meta672_down)
acc_6100_down <- length(intersect_6100_down)/nrow(meta6100_down)

acc_955_down <- length(intersect_955_down)/nrow(meta955_down)
acc_972_down <- length(intersect_972_down)/nrow(meta972_down)
acc_9100_down <- length(intersect_9100_down)/nrow(meta9100_down)

# acc_DE (up or down)
acc_355_DE <- length(intersect_355_DE)/nrow(meta355_DE)
acc_372_DE <- length(intersect_372_DE)/nrow(meta372_DE)
acc_3100_DE <- length(intersect_3100_DE)/nrow(meta3100_DE)

acc_655_DE <- length(intersect_655_DE)/nrow(meta655_DE)
acc_672_DE <- length(intersect_672_DE)/nrow(meta672_DE)
acc_6100_DE <- length(intersect_6100_DE)/nrow(meta6100_DE)

acc_955_DE <- length(intersect_955_DE)/nrow(meta955_DE)
acc_972_DE <- length(intersect_972_DE)/nrow(meta972_DE)
acc_9100_DE <- length(intersect_9100_DE)/nrow(meta9100_DE)
```

## Create data frame of EdgeR
```{r}
size_sample <- c("3_500_500", "3_750_250", "3_1000_0", 
                 "6_500_500", "6_750_250", "6_1000_0", 
                 "9_500_500", "9_750_250", "9_1000_0")
num_intersec_up <- c(length(intersect_355_up), length(intersect_372_up), length(intersect_3100_up),
                  length(intersect_655_up), length(intersect_672_up), length(intersect_6100_up),
                  length(intersect_955_up), length(intersect_972_up), length(intersect_9100_up))
num_meta_up <- c(nrow(meta355_up), nrow(meta372_up), nrow(meta3100_up),
                 nrow(meta655_up), nrow(meta672_up), nrow(meta6100_up),
                 nrow(meta955_up), nrow(meta972_up), nrow(meta9100_up))

num_intersec_down <- c(length(intersect_355_down), length(intersect_372_down), length(intersect_3100_down),
                  length(intersect_655_down), length(intersect_672_down), length(intersect_6100_down),
                  length(intersect_955_down), length(intersect_972_down), length(intersect_9100_down))
num_meta_down <- c(nrow(meta355_down), nrow(meta372_down), nrow(meta3100_down),
                 nrow(meta655_down), nrow(meta672_down), nrow(meta6100_down),
                 nrow(meta955_down), nrow(meta972_down), nrow(meta9100_down))

num_intersec_DE <- c(length(intersect_355_DE), length(intersect_372_DE), length(intersect_3100_DE),
                  length(intersect_655_DE), length(intersect_672_DE), length(intersect_6100_DE),
                  length(intersect_955_DE), length(intersect_972_DE), length(intersect_9100_DE))
num_meta_DE <- c(nrow(meta355_DE), nrow(meta372_DE), nrow(meta3100_DE),
                 nrow(meta655_DE), nrow(meta672_DE), nrow(meta6100_DE),
                 nrow(meta955_DE), nrow(meta972_DE), nrow(meta9100_DE))

acc_up <- round(c(acc_355_up, acc_372_up, acc_3100_up, 
            acc_655_up, acc_672_up, acc_6100_up, 
            acc_955_up, acc_972_up, acc_9100_up), digits = 3)
acc_down <- c(acc_355_down, acc_372_down, "undefined", 
            acc_655_down, acc_672_down, "undefined", 
            acc_955_down, acc_972_down, "undefined")
acc_DE <- round(c(acc_355_DE, acc_372_DE, acc_3100_DE, 
            acc_655_DE, acc_672_DE, acc_6100_DE, 
            acc_955_DE, acc_972_DE, acc_9100_DE), digits =3)

edgeR_acc <- data.frame(size_sample, num_intersec_up, num_meta_up, acc_up, num_intersec_down, num_meta_down, acc_down, num_intersec_DE, num_meta_DE, acc_DE)
edgeR_acc


```

```{r}
# export .csv files
write.csv(edgeR_acc,"Accuracy_EdgeR_Vs_meta.csv", row.names = FALSE)
```

